# Priors as Hidden Mixtures {.unnumbered}

```{r, message=F, warning=F}
library(cmdstanr)
```

It is adapted from [Unit 7: tasmixture.odc](../odc_files/unit7/tasmixture.odc).

## Problem Statement

We are comparing the Student T likelihood with Normal prior vs Normal likelihood with Gamma prior.

### Model

```{r file_location, include=FALSE}
tas_mixture <- file.path('stan_models/tas_mixture.stan')
```

```{r, warning=FALSE, comment=NA, error=FALSE, message=FALSE}
mod <- cmdstan_model(tas_mixture)
```

now that we have compiled our stan file, we can print out the model that we will use for this:

```{r print_model}
mod$print()
```

### Sampling

Let us prepare the data to be passed into sample

```{r}
input_data <- list(
  df = 6,
  mu0 = 6,
  tau1 = 10,
  tau = 0.4,
  X = 10,
  Y = 10
)
```

Now that we have the data to pass into our sampling, let's proceed.

```{r sampling, message=F, warning=F, results='hide'}
fit <- mod$sample(
  data = input_data,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  refresh = 500,   # print update every 500 iterations
  iter_warmup = 1000,
  iter_sampling = 5000
)
```

```{r, comment=NA}
fit$summary()
```

These results are very similar to PyMC results.
